---
title: "Classifying Covid-19 Test Results"
subtitle: "A Tidymodels Workflow"
author: "Itamar Caspi"
date: '(updated: `r Sys.Date()`)'
output:
  html_document:
    code_folding: show
    highlight: haddock
    keep_md: no
    theme: journal
    toc: yes
    toc_depth: 4
    toc_float: yes
---

# Load packages

Load packages
```{r pacman, message=FALSE, warning=FALSE, eval=TRUE}
if (!require("pacman")) install.packages("pacman")

pacman::p_load(
  tidyverse,   # for data wrangling and visualization
  tidymodels,  # for data modeling
  vip,         # for variable importance
  here,        # for referencing files and folders
  readxl       # for reading xlsx files
)
```

and set seed for replication
```{r}
set.seed(1203)
```

# Read data

```{r read_data}
covid_raw <- 
  here("06-classification/data", "corona_tested_individuals_ver_002.xlsx") %>% 
  read_xlsx()
```


# Initial preprocessing

Replace all `NULL` values with `NA`'s
```{r}
is.na(covid_raw[,-1]) <- covid_raw[,-1] == "NULL"
```


```{r}
covid <- 
  covid_raw %>%
  select(-test_date, -test_indication) %>% 
  filter(corona_result != "אחר") %>% 
  drop_na() %>%
  mutate(
    corona_result = if_else(corona_result == "שלילי", "negative", "positive"),
    gender = if_else(gender == "זכר", "male", "female")
  ) %>% 
  mutate_all(as_factor)
  
```

# Save processed data
```{r save}
write.csv(
  covid,
  file = here("06-classification/data","covid_proc.csv"),
  row.names = FALSE
  )
```

# Split

Train-test split
```{r}
covid_split <- covid       %>% initial_split(prop = 0.5)
covid_train <- covid_split %>% training()
covid_test  <- covid_split %>% testing()
```

Cross-validation splits
```{r}
covid_folds <- covid_train %>% vfold_cv(v = 5, strata = corona_result)
```

# Build a Model

```{r}
logit_model <- 
  logistic_reg() %>%
  set_engine("glmnet") %>% 
  set_mode("classification") %>%
  set_args(penalty = tune(), mixture = 1)
```

# Create the Recipe

```{r}
covid_rec <- 
  recipe(corona_result ~ ., data = covid_train) %>% 
  step_dummy(all_nominal(), -corona_result, one_hot = TRUE) %>% 
  step_interact(~ all_predictors():all_predictors()) %>%
  step_normalize(all_predictors()) %>% 
  step_zv(all_predictors()) 
```


# Create the Workflow

```{r}
logit_wfl <- 
  workflow() %>% 
  add_recipe(covid_rec) %>% 
  add_model(logit_model)
```

# Train and Tune the Model

```{r}
roc_only <- metric_set(roc_auc)

logit_result <- 
  logit_wfl %>% 
  tune_grid(
    resamples = covid_folds,
    control = control_grid(save_pred = TRUE),
    metrics = roc_only
  )
```

# Plot Cross-validation results
```{r}
logit_result %>% 
  collect_metrics() %>% 
  ggplot(aes(x = penalty, y = mean)) + 
  geom_point() + 
  geom_line() + 
  ylab("Area under the ROC Curve") +
  scale_x_log10(labels = scales::label_number())

```


show best results
```{r}
logit_result %>% 
  show_best(metric = "roc_auc", n = 10)
```

## Set best lambda

Two options: lambda that minimizes `roc_auc`, and the 1 standard error rule of thumb:
```{r}
lambda_min <- logit_result %>% 
  select_best(
    metric = "roc_auc"
  )

lambda_1se <- logit_result %>% 
  select_by_one_std_err(
    metric = "roc_auc",
    desc(penalty)
  ) %>% 
  select(penalty)
```

## Last fit

Final workflow
```{r}
logit_wfl_final <- 
  logit_wfl %>%
  finalize_workflow(lambda_1se)
```

Last fit on the test set
```{r}
logit_last_fit <- 
  logit_wfl_final %>% 
  last_fit(covid_split)
```

# ROC AUC on the train and test sets

```{r}
logit_wfl_final %>%
  fit(data = covid_train) %>% 
  predict(new_data = covid_train, type = "prob") %>% 
  roc_curve(covid_train$corona_result, .pred_negative) %>% 
  autoplot() +
  labs(title = "Training set AUC")
```

```{r}
logit_last_fit %>% 
  collect_predictions() %>% 
  roc_curve(corona_result, .pred_negative) %>% 
  autoplot() +
  labs(title = "Test set AUC")
```

```{r}
logit_last_fit %>% 
  collect_metrics() %>% 
  filter(.metric == "roc_auc")
```

# Variable importance
```{r}
logit_last_fit %>% 
  pluck(".workflow", 1) %>%   
  pull_workflow_fit() %>% 
  vip(num_features = 10)
```
